---
title: "Floral scent of *Schiedea kaalae x hookeri* hybrids"
author: "[John Powers](http://sites.uci.edu/powers/)"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: no
    lib_dir: libs
    code_folding: hide
    toc: yes
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---
<style type="text/css">
.main-container { max-width: 1600px; margin-left: 0; margin-right: auto; }
img{ max-width:200%; height: auto; }
td, th { padding : 6px }
</style>

```{r setup, include=FALSE}
library(tidyverse)
library(vegan)
library(RColorBrewer)
library(knitr)
library(emmeans)
library(broom)
library(ggpubr)
library(ggvegan)
library(ggordiplots)
library(seriation)
library(dendsort)
library(pheatmap)
library(viridis)
library(grid)
#also uses multcomp

knitr::opts_chunk$set(comment="", cache=T, warning = F, message = F, 
                      fig.path = "hybrids_cut_plots/", dev="svglite", dev.args=list(fix_text_size=FALSE))
```

```{r figures, eval=FALSE}
figs <- list.files("./hybrids_cut_plots", pattern=".svg")
walk(figs, ~system(paste0("inkscape --export-dpi 150 -o ./hybrids_cut_plots/png/", str_replace(.x, ".svg",".png"), " ./hybrids_cut_plots/", .x)))
walk(figs, ~system(paste0("inkscape -o ./hybrids_cut_plots/eps/", str_replace(.x, ".svg",".eps"), " ./hybrids_cut_plots/", .x)))
```

```{r sdata}
greenhouse <- matrix(c(-117.8475746, 33.6472914), nrow=1) #latlong of UCI greenhouse

s <- read_tsv("data/UCI_GCMS/Schiedea Volatiles Sampling - Samples.tsv") %>% 
  drop_na(Start, Stop) %>% #2 bad samples with no Start, 1 missing Stop in 2019
  nest(.by=SampleDate) %>% #sunset function is slow, run on unique dates
  mutate(sunset = map_vec(SampleDate, 
                          ~maptools::sunriset(dateTime=ymd_hms(paste(.x, "13:00:00")),
                                              direction="sunset", crds=greenhouse, POSIXct.out=T)$time),
         solarnoon = map_vec(SampleDate, 
                          ~maptools::solarnoon(dateTime=ymd_hms(paste(.x, "13:00:00")),
                                              crds=greenhouse, POSIXct.out=T)$time)) %>% unnest(data) %>% 
  mutate(StartSunset = difftime(ymd_hms(paste(SampleDate, Start)), sunset, units="hours"),
         StopSunset =  difftime(ymd_hms(paste(SampleDate, Stop)), sunset, units="hours"),
         StartNoon =   difftime(ymd_hms(paste(SampleDate, Start)), solarnoon, units="hours"))

CASsubs <- read_tsv("data/UCI_GCMS/CASsubs3.csv") %>% select(CAS, newCAS) %>% deframe()

data <- read_tsv("data/UCI_GCMS/svhyb_ambient_blanks.csv") %>% 
  distinct() %>% #generated some reports twice
  select(FileName:`Base Peak`, Weighted, Reverse) %>% 
  rename(Path = FileName, Fraction = Amount, Match = Weighted) %>% 
  mutate(Path = str_sub(Path, 52, -5), #remove base directory and file extension
         SampleDate = str_extract(Path, "\\\\(.*)", group=1), #file name after folder
         Sample = if_else(str_sub(SampleDate, 1, 5)=="BLANK", SampleDate, 
          str_extract(SampleDate, "(.*?)_.*", group=1) %>% #drop GC index
            #pad sample number to 3 digits with zeros
            str_replace("(\\d+)", function(x) formatC(as.integer(x), width=3, format="d", flag="0")) %>% 
            str_replace("^KAHO", "SKAHO")),
         CAS = str_replace(CAS, "(\\d{2,7})(\\d{2})(\\d{1})$","\\1-\\2-\\3") %>% #add dashes
           recode(!!!CASsubs),#manual identifications
         Name = str_remove(Name, "^>"),
         across(c(SampleDate, Path, Model, Sample, Name, CAS), factor),
         across(c(Fraction, Purity, `Min. Abund.`), ~ as.numeric(str_remove(.x, "%"))/100)) %>% 
  #AMDIS Base Peak only has the most common ion integrated
  #instead, multiply fractions by total peak areas from sum_chromatograms.R
  left_join(read_tsv("data/UCI_GCMS/svhyb_TICsums.csv") %>% 
              mutate(SampleDate = str_extract(file, "(.*).CDF", group=1))) %>% 
  mutate(Area = Fraction * sumTIC)

chems.pc <- read_tsv("data/UCI_GCMS/chemspc3.csv") %>% select(CAS, IUPAC.Name) %>% mutate(IUPAC.Name = str_replace(IUPAC.Name, "~\\{(.*?)\\}", "\\1"))
```

# Filtering

```{r ng_hr}
#Load filtered scent data (includes octenol and hexenol) and metadata
# load("./data/UCI_GCMS/sdata3_hexoct.Rdata")
# TODO some red flags in the Notes column - possible wrong IDs
# sv$spec <- spec
# save(sv, vol, file = "./data/UCI_GCMS/sdata3_hexoct_svvol.Rdata")
load("./data/UCI_GCMS/sdata3_hexoct_svvol.Rdata")

#convert to nanograms per hour using standards for each compound class
vocs <- colnames(vol)
chemsf <- tibble(shortname = vocs) %>%# two VOCs got renamed since table was made 
  bind_cols(read_tsv("data/UCI_GCMS/chemspc3_classes.csv")) %>% 
  left_join(read_tsv("data/UCI_GCMS/multipliers.csv"))
vol <- sweep(vol, 2, chemsf$ngPerArea, FUN= `*`)
```

# Sample inventory

```{r sv}
eve <- -2.5 #hours before sunset
hybcolr <- c(`S. kaalae` = "#31A354", `K x H` = "#FD8D3C", `H x K` = "#A65628", `S. hookeri` = "#756BB1")
hyblabs <- c(expression(italic("S. kaalae")),"K x H", "H x K", expression(italic("S. hookeri")))
renamepops <- c(`879WKG` = "879", WK = "794", WK = "866", WK = "891", WK = "899", #lump Waianae Kai populations
                `3587WP` = "3587", `892WKG` = "892", `904WPG` = "904")
kahopops <- unique(names(renamepops))
hybpopsp <- c(K = "3587WP", H = "879WKG", K = "881KHV", K = "892WKG", K = "904WPG",
              HH = "HH", HK = "HK", KH = "KH", KK = "KK", H = "WK")

#Get maternal and paternal populations of crosses
kahomp <- sv %>% rownames_to_column() %>% filter(Species == "kaho") %>% select(rowname, Plant) %>% 
  separate(Plant, into = c("Maternal", "Paternal"), sep=" x ") %>% 
  mutate(across(ends_with("aternal"), ~str_split_i(.x, "-", 1) %>% 
                  fct_recode(!!!renamepops) %>% fct_relevel(kahopops)))

#Add columns to metadata describing the plant
svhyb <- sv %>% rownames_to_column() %>% left_join(kahomp) %>% 
  mutate(specp = case_match(Population, "KK" ~ "KAAL", "HH" ~ "HOOK", .default=spec) %>% 
           fct_relevel(c("HOOK","KAHO","KAAL")), #Split cross offspring (spec=KAHO) into parents and hybrids 
         Population2 = factor(if_else(Species == "kaho", if_else(Maternal==Paternal, Maternal, "Interpop"), Population)),
         SpeciesR = if_else(specp=="KAHO", Population, specp) %>% 
           factor(levels=c("KAAL","KH","HK","HOOK")) %>% 
           fct_recode("S. kaalae"="KAAL", "K x H"="KH", "H x K"="HK", "S. hookeri"="HOOK"),
         DN = factor(if_else(StartSunset > eve, "Night", "Day")),
         Inflo = if_else(is.na(Inflo), as.character(1:n()), Inflo),
         Diurnal = if_else(is.na(Diurnal), as.character(1:n()), Diurnal),
         plants =          paste(Population, Plant, sep="-"),
         plantsi =         paste(Population, Plant, Cutting, sep="-"),
         plantsinflo =     paste(Population, Plant, Cutting, Inflo, sep="-"),
         plantsinflodate = paste(Population, Plant, Cutting, Inflo, SampleDate, sep="-"),
         parents = if_else(str_detect(plants, fixed(" x ")), as.character(1:n()), plants),
         ftot = rowSums(vol) / Flrs) %>% 
  arrange(specp, Population)

svhybn <- filter(svhyb, DN == "Night")
#"parents" set the crosses to an integer sequence to force them not to group 
#TODO this unfortunately splits resamples of the sample cross plant into multiple "parents"
#TODO also, NMDS below implies these are different cross parents but they are really just the original plant in the parent generation that was used for cuttings
```

```{r plantmeans}
volhybn <- vol[svhybn$rowname,] #emission rate, nanograms per hour
pavolhybn <- decostand(volhybn, method = "pa") #presence/absence
fvolhybn <- volhybn / svhybn$Flrs #emission rate per flower

#first take the average for each infloresence
#Inflos are marked by a letter but those are reused arbitrarily across SampleDates
svhybn.inflo <- bind_cols(svhybn, fvolhybn) %>% 
 group_by(SpeciesR, Species, Population, Maternal, Paternal, plants, plantsi, Inflo, SampleDate) %>% 
  summarize(across(all_of(vocs), mean), .groups = "drop")
fvolhybn.inflo <- svhybn.inflo[,vocs]

#now average infloresences by plant
#TODO temporarily got rid of parents since it was splitting multiple inflos on same plant
svhybn.plants <- svhybn.inflo %>% 
 group_by(SpeciesR, Species, Population, Maternal, Paternal, plants, plantsi) %>% 
  summarize(across(all_of(vocs), mean), .groups = "drop")
fvolhybn.plants <- svhybn.plants[,vocs]
```

## Samples by cross

There are also four 881KHV samples (not used to make crosses since from the Ko'olaus), and two samples without populations (an HH and a KH).

```{r crosses}
inventory.cross <- svhybn %>% filter(Species == "kaho") %>%
  drop_na(Maternal, Paternal) %>% #two samples where parent populations are unknown
  with(table(Maternal, Paternal)) 
inventory.all <- inventory.cross + diag(table(svhybn$Population)[rownames(inventory.cross)])
kable(inventory.all, caption="evening samples of crosses and parents")

#how many maternal and paternal parent plants?
mp <- svhybn %>% filter(Species =="kaho") %>% select(Plant) %>% mutate(Plant=as.character(Plant)) %>%
  separate(Plant, sep=" x ", into = c("Maternal", "Paternal")) %>% 
  separate(Maternal, sep="-", into = c("MomPop", "MomPlant"), remove=F, extra="merge") %>% 
  separate(Paternal, sep="-", into = c("DadPop", "DadPlant"), remove=F, extra="merge")
mp %>% group_by(MomPop) %>% summarize(moms=n_distinct(MomPlant)) %>% kable(caption="maternal plants")
mp %>% group_by(DadPop) %>% summarize(moms=n_distinct(DadPlant)) %>% kable(caption="paternal plants")
```

## Resampled plants

```{r duplicates}
svhybn.inflo %>% count(SpeciesR, plantsi) %>% filter(n>1) %>% kable(caption = "number of infloresences per plant")
```

## Evening plants by cross

remember to add the 3 881KHV plants (not used to make crosses since from the Ko'olaus)

there are also two plants from crosses without populations (HH and KH) - added these to 879x879 and 892xWK in the inventory

```{r crosses_plants}
plants.cross <- svhybn.plants %>% filter(Species == "kaho") %>%
  drop_na(Maternal, Paternal) %>% #two samples where parent populations are unknown
  with(table(Maternal, Paternal)) 
kable(plants.cross, caption="cross plants")

plants.parent <- diag(table(svhybn.plants$Population)[rownames(plants.cross)])
rownames(plants.parent) <- rownames(plants.cross); colnames(plants.parent) <- colnames(plants.cross)
kable(plants.parent, caption="parent plants")
```

# Sampling times

```{r times}
print("night pumping times, hrs:")
summary(as.numeric(as.difftime(format.POSIXct(svhybn$SampleDateStart, tz="GMT+8", format="%H:%M"),format="%H:%M"))) #convert PDT to PST when needed

print("night pumping times, hr since sunset:")
summary(as.numeric(svhybn$StartSunset))
```

# Heatmap

```{r heatmap, dev="png", dev.args=list(), fig.width=10.5, fig.height=7.6}
class_subs <- c("Irregular terpene"="Other terpenoid", "Monoterpene"="Monoterpenoid")
classcolr <- c(Aliphatic="#BC0060",Benzenoid="#027B8C",Monoterpenoid="#5d9400",`Other terpenoid`="#E56E00", Sesquiterpene="#000000")

#1 dimensional NMDS for ordering samples in heatmap
set.seed(1)
mdshybn1 <- metaMDS(decostand(fvolhybn.plants, method="hellinger"),k=1,try=200, autotransform=F, trace=0)
srt <- order(svhybn.plants$SpeciesR, -mdshybn1$points, decreasing=T)

#Include volatiles found in at least this many samples
mincount <- 15

callback = function(hc, mat){
    sv = svd(t(mat))$v[,1]
    dend = reorder(as.dendrogram(hc), wts = sv)
    as.hclust(dend)
}
heat.mat <- as.matrix(t(fvolhybn.plants[,colSums(pavolhybn)>mincount][srt,]))^(1/3)
colnames(heat.mat) <- rownames(fvolhybn.plants)[srt]
ph <- pheatmap(heat.mat, 
         cluster_cols=F, show_colnames=F,
         clustering_method="mcquitty", clustering_distance_rows="correlation",
         clustering_callback = function(hc, ...){dendsort(hc, type="average")},
         scale="none", color=inferno(512),
         annotation_col = data.frame(Species=as.integer(svhybn.plants$SpeciesR)[srt], row.names=rownames(fvolhybn.plants)[srt]), 
         annotation_colors = list(Species=hybcolr, Class=classcolr), annotation_names_col = F,
         gaps_col = which(as.logical(diff(as.integer(svhybn.plants$SpeciesR)[srt]))),
         annotation_row = data.frame(Class=chemsf$Class %>% str_remove("s$") %>% recode(!!!class_subs) %>% factor(), 
                                     row.names = chemsf$shortname), annotation_names_row = F,
   cellwidth = 3.5, cellheight = 14, fontsize = 10, border_color = NA, legend=F, legend_breaks=NA, annotation_legend=F, cutree_rows=5
)
downViewport("col_annotation.3-3-3-3")
middles <- enframe(svhybn.plants$SpeciesR[srt]) %>% mutate(name=(name-0.5)/max(name-0.5)) %>% summarize(name=mean(name), .by=value) %>% pull(name)
grid.text(rev(levels(svhyb$SpeciesR))[c(1,4)], x=middles[c(1,4)],y=0.57, gp=gpar(fontsize=10, col="white", fontface=4))
grid.text(rev(levels(svhyb$SpeciesR))[c(2,3)], x=middles[c(2,3)],y=0.57, gp=gpar(fontsize=10, col="white", fontface=2))
```

# Boxplots of each volatile

```{r boxplots, fig.width=9, fig.height=12}
diversity_vars <- c(ftot = "Total volatile emissions (ng/flr/hr)", 
                    num_compounds="Number of compounds", 
                    shannon="Shannon diversity index")
svhybn.plants$ftot <- rowSums(fvolhybn.plants)
svhybn.plants$num_compounds <- rowSums(decostand(fvolhybn.plants, method="pa"))
svhybn.plants$shannon <- diversity(fvolhybn.plants, index="shannon")

multiline <- c("DMNT"="(3E)-4,8-dimethylnona-1,3,7-triene",
               "2,2,6-trimethylcyclohexane-\n1,4-dione"="2,2,6-trimethylcyclohexane-1,4-dione",
               "4-hydroxy-2,6,6-trimethyl-\n3-oxocyclohexene-\n1-carbaldehyde"=
                 "4-hydroxy-2,6,6-trimethyl-3-oxocyclohexene-1-carbaldehyde")

minprop <- 0.25 #compound must occur in greater than this proportion of evening samples
fvolhybn.top <- fvolhybn.plants %>% 
  select(where(~sum(.x>0) > length(.x) * minprop)) %>% 
  bind_cols(select(svhybn.plants, Species=SpeciesR, all_of(names(diversity_vars)))) %>% 
  pivot_longer(!Species, names_to="variable") %>% 
  mutate(variable = fct_recode(variable, !!!multiline))

#compare mean of hybrid cross directions to mean of parent species
#compare hybrid cross directions to each other
tests.glht <- fvolhybn.top %>% group_by(variable) %>% nest() %>% 
  mutate(model = map(data, ~lm(value ~ Species, data=.x)),
         emm = map(model, ~tidy(summary(emmeans(.x, specs="Species")))),
         test.direction = map(model, ~tidy(multcomp::glht(.x, linfct = multcomp::mcp(Species=c("`K x H` - `H x K` == 0"))))),
         test.parmean = map(model, ~tidy(multcomp::glht(.x, linfct = multcomp::mcp(Species=c("0.5 * `K x H` + 0.5 * `H x K` - 0.5 * `S. kaalae` - 0.5 * `S. hookeri`== 0"))))))

tests.emm <- tests.glht %>% mutate(maximum = map_dbl(data, ~max(.x$value))) %>% 
  select(variable, emm, maximum) %>% unnest(emm)
tests.index <- tests.emm %>% 
  select(variable, Species, maximum, estimate) %>% pivot_wider(names_from=Species, values_from=estimate) %>% 
  mutate(par.mean = (`S. hookeri` + `S. kaalae`)/2,
         hyb.mean = (`K x H`+`H x K`)/2,
         hyb.index = if_else(`S. kaalae`> `S. hookeri`, 
                             (hyb.mean -`S. hookeri`)/ `S. kaalae`, 
                             (hyb.mean -`S. kaalae`)/ `S. hookeri`))
tests.direction <- tests.glht %>% select(variable, test.direction) %>% unnest(test.direction) %>% 
  select(variable, direction.p.value = adj.p.value)
tests.parmean <- tests.glht %>% select(variable, test.parmean) %>% unnest(test.parmean) %>% 
  select(variable, parmean.p.value = adj.p.value)
boxplot.order <- tests.index %>% arrange(-hyb.index) %>% pull(variable)
tests.all <- tests.index %>% left_join(tests.direction) %>% left_join(tests.parmean) %>%
  mutate(variable=factor(variable, levels=boxplot.order))

tests.vols <- tests.all %>% filter(!variable %in% names(diversity_vars)) %>% 
    mutate(across(contains("p.value"), ~p.adjust(.x, method="fdr")))
tests.vols %>% select(-c(maximum, hyb.index)) %>% mutate(variable=str_remove_all(variable, "\n")) %>% 
  kable(caption = "tests of total emissions and diversity", digits=3)
emm.vols <- tests.emm %>% filter(!variable %in% names(diversity_vars)) 

signif.num <- function(x) {
   cut(x, breaks = c(0, 0.001, 0.01, 0.05, 1), 
           labels = c("***", "**", "*", " "), include.lowest = T)
}
fvolhybn.top %>% mutate(variable=factor(variable, levels=boxplot.order)) %>% 
  filter(!variable %in% names(diversity_vars)) %>% 
  ggplot() +  geom_boxplot(aes(x=Species, y=value, color=Species), width=0.5, outlier.size=1) + 
  geom_text(data=tests.vols, aes(label=signif.num(parmean.p.value), x=2.5, y=pmax(par.mean,hyb.mean)*1.5)) + 
  geom_point(data=emm.vols, aes(x=Species, y=pmax(estimate,0), color=Species), shape=4, size=2)+
  geom_point(data=tests.vols, aes(x=2.5, y=par.mean, shape="parmean"), size=2)+
  geom_point(data=tests.vols, aes(x=2.5, y=hyb.mean, shape="hybmean"), size=2)+
  scale_shape_manual("Mean",values=c(parmean=3, hybmean=5), labels=c(parmean="Species", hybmean="Hybrids"))+
  facet_wrap(vars(variable), scales="free_y", ncol=4, labeller = labeller(variable = label_wrap_gen(18))) +
  scale_color_manual("", values=hybcolr, labels=hyblabs) +
  scale_y_sqrt("Emission rate (ng/flower/hr)", limits=c(0,NA)) + theme_minimal() + 
  theme(axis.text.x=element_blank(), axis.title.x=element_blank(), 
        panel.grid.major.x=element_blank(), panel.grid.minor.y=element_blank(), legend.position = "top")
```

## Hybrid additivity vs. differences between parents 

```{r hyb_combo, fig.height=7, fig.width=7}
library(ggrepel)
vol_floor <- 1e-5 #avoid negative estimates, 10x less than smallest positive estimate
tests.vols %>%  mutate(lfc_species = log2(pmax(`S. kaalae`,vol_floor)/pmax(`S. hookeri`,vol_floor)), 
                       lfc_hybrids = log2(hyb.mean/par.mean)) %>% ungroup() %>% 
  left_join(chemsf %>% mutate(variable=fct_recode(Name, !!!multiline))) %>% 
  mutate(Class = Class %>% str_remove("s$") %>% recode(!!!class_subs) %>% factor()) %>% 
  ggplot(aes(x=lfc_species, y=lfc_hybrids, color=Class)) + 
  geom_vline(xintercept=0, color="grey40") + geom_hline(yintercept=0, color="grey40") + 
  geom_point(aes(shape = parmean.p.value < 0.05), size=2.5) + scale_shape_manual(values=c(1,19), guide="none")+
  geom_text_repel(aes(label=variable), size=3.5, min.segment.length=5, show.legend = F)  +
  scale_color_manual(values=classcolr) +
  scale_x_continuous(expand=expansion(c(0.1, 0.25)))+ 
  theme_minimal() + theme(panel.grid.minor = element_blank(), legend.position = "top") +
  labs(x="log2 fold change in emissions between parents (S. kaalae / S. hookeri)",
       y="log2 fold change in emissions between hybrids and mean of parents",color="")
```


# Total emissions and diversity

```{r diversity, fig.height=4, fig.width=6.5}
tests.diversity <- tests.all %>% filter(variable %in% names(diversity_vars)) %>% 
    mutate(across(contains("p.value"), ~p.adjust(.x, method="fdr"))) 
tests.diversity %>% select(-c(maximum, hyb.index)) %>% kable(caption = "tests of total emissions and diversity", digits=3)
emm.diversity <- tests.emm %>% filter(variable %in% names(diversity_vars)) 

fvolhybn.top %>% filter(variable %in% names(diversity_vars)) %>% 
  ggplot() +  geom_boxplot(aes(x=Species, y=value, color=Species), width=0.5, outlier.size=1) + 
  geom_text(data=tests.diversity, aes(label=signif.num(parmean.p.value), x=2.5, y=pmax(par.mean,hyb.mean)*1.1)) + 
  geom_point(data=emm.diversity, aes(x=Species, y=pmax(estimate,0), color=Species), shape=4, size=2)+
  geom_point(data=tests.diversity, aes(x=2.5, y=par.mean, shape="parmean"), size=2)+
  geom_point(data=tests.diversity, aes(x=2.5, y=hyb.mean, shape="hybmean"), size=2)+
  scale_shape_manual("Mean",values=c(parmean=3, hybmean=5), labels=c(parmean="Species", hybmean="Hybrids"))+
  facet_wrap(vars(variable), scales="free_y", ncol=4, labeller = as_labeller(diversity_vars)) +
  scale_color_manual("", values=hybcolr, labels=hyblabs) + 
  scale_y_continuous("", limits=c(0,NA)) + theme_minimal() + 
  theme(axis.text.x=element_blank(), axis.title.x=element_blank(), 
        panel.grid.major.x=element_blank(), panel.grid.minor.y=element_blank(), legend.position = "top")
```

# NMDS ordination

## Each sample (includes resamples)

```{r 2dnmdsnight, fig.height=8, fig.width=10.5}
set.seed(1)
mdshybn <- metaMDS(decostand(fvolhybn, method="hellinger"),k=2,try=200, autotransform=F, trace=0)
mdshybn

hullcross <- gg_ordiplot(mdshybn, groups = svhybn$plants, plot=F)$df_hull
hullparents <- gg_ordiplot(mdshybn, groups = svhybn$parents, plot=F)$df_hull
hullplantsi <- gg_ordiplot(mdshybn, groups = svhybn$plantsi, plot=F)$df_hull
hullplantsinflo <- gg_ordiplot(mdshybn, groups = svhybn$plantsinflo, plot=F)$df_hull
hullplantsinflodate <- gg_ordiplot(mdshybn, groups = svhybn$plantsinflodate, plot=F)$df_hull

obj <- fortify(mdshybn)
obj$Sample <- NA; obj$Sample[obj$score=="sites"] <-  svhybn$Sample
obj <- obj %>% left_join(svhybn %>% dplyr::select(c(Sample, SpeciesR, plants, parents, plantsi, plantsinflo, plantsinflodate, StartSunset))) %>% mutate(StartSunset = as.numeric(StartSunset))
obj$occur <- NA; obj$occur[obj$score=="species"] <- colSums(pavolhybn)
obj$ftotal <- NA; obj$ftotal[obj$score=="species"] <- colSums(fvolhybn)
obj <- obj %>% arrange(NMDS2, NMDS1)

hybnplot <- ggplot(obj, aes(x=-NMDS1, y=NMDS2)) + xlab("NMDS1") +
    geom_path(data=filter(obj, score=="sites"),aes(group=plants, color="cross"), linewidth=1) +
    geom_path(data=filter(obj, score=="sites"),aes(group=parents, color="parents"), linewidth=1) +
    geom_path(data=filter(obj, score=="sites"),aes(group=plantsi, color="plantsi"), linewidth=1) +
    geom_path(data=filter(obj, score=="sites"),aes(group=plantsinflo, color="plantsinflo"), linewidth=1) +
    geom_path(data=filter(obj, score=="sites"),aes(group=plantsinflodate, color="plantsinflodate"), linewidth=1) +
  scale_color_manual("",breaks=c("plantsinflodate", "plantsinflo", "plantsi", "parents", "cross"),
                     labels=   c("Bag",             "Infloresence","Plant",   "Clone",   "Cross"),
                     values=c(cross="grey80", parents="grey50", plantsi="grey20",
                              plantsinflo="goldenrod",plantsinflodate="#E41A1C",sites="white",species=NA)) +
<<<<<<< HEAD
      coord_fixed(xlim=range(obj$NMDS1[obj$score=="sites"])+c(-0.35,0.45), 
=======
      coord_fixed(xlim=range(obj$NMDS1[obj$score=="sites"])+c(-0.42,0.45), 
>>>>>>> 27b13227da49979dfa7b2b0cafc285e7cc4f0a95
                  ylim=range(obj$NMDS2[obj$score=="sites"])) + theme_pubr()+
    scale_x_continuous(expand=c(0,0)) +
    theme(legend.text = element_text(size=13)) + 
    guides(fill = guide_legend(override.aes = list(size=6)), color=guide_legend(override.aes=list(size=4)))

(hybplotn.spec <- hybnplot + 
        scale_fill_manual("", values=hybcolr, labels=hyblabs,  na.translate=F) +
  geom_point(data=obj[obj$score=="sites",], aes(fill=SpeciesR), shape=21, size=3, color="white") +
    geom_text(data=obj[obj$score=="species" & obj$occur>nrow(volhybn)*minprop,], aes(label=label), color="black", size=3.8, fontface=2) ) 
```

## Mean of each plant

```{r 2dnmdsnightmean, fig.height=8, fig.width=10.5}
set.seed(1)
mdshybn.mean <- metaMDS(decostand(fvolhybn.plants, method="hellinger"),k=2,try=100, autotransform=F, trace=0)
mdshybn.mean

obj.mean <- fortify(mdshybn.mean)
obj.mean$SpeciesR <- factor(NA,levels=levels(svhybn.plants$SpeciesR)); obj.mean$SpeciesR[obj.mean$score=="sites"] <-  svhybn.plants$SpeciesR
obj.mean$plants <- factor(NA,levels=unique(svhybn.plants$plants)); obj.mean$plants[obj.mean$score=="sites"] <-  svhybn.plants$plants
#obj.mean$parents <- factor(NA,levels=levels(svhybn.plants$parents)); obj.mean$parents[obj.mean$score=="sites"] <-  svhybn.plants$parents
obj.mean$occur <- NA; obj.mean$occur[obj.mean$score=="species"] <- colSums(pavolhybn)
obj.mean$ftotal <- NA; obj.mean$ftotal[obj.mean$score=="species"] <- colSums(fvolhybn.plants)
obj.mean <- obj.mean %>% arrange(NMDS2, NMDS1)

(hybplotn.mean.spec <- 
    ggplot(obj.mean, aes(x=NMDS1, y=NMDS2)) + 
    coord_fixed(xlim=range(obj.mean$NMDS1[obj.mean$score=="sites"])+c(-0.55,0.1), 
                ylim=range(obj.mean$NMDS2[obj.mean$score=="sites"])) + 
    theme_pubr()+
    scale_x_continuous(expand=c(0,0))+
    theme(legend.text = element_text(size=13)) + 
    guides(fill = guide_legend(override.aes = list(size=6)), color=guide_legend(override.aes=list(size=4)))+
    scale_fill_manual("", values=hybcolr, labels=hyblabs,  na.translate=F) +
      scale_color_manual("",breaks=c("cross","parents"), labels=c("Cross","Clone"), values=c(cross="grey80")) +
    geom_path(data=obj.mean[obj.mean$score=="sites",],aes(group=plants, color="cross"), size=1) +
    #geom_path(data=obj.mean[obj.mean$score=="sites",],aes(group=parents, color="parents"), size=1) +
    geom_point(data=obj.mean[obj.mean$score=="sites",], aes(fill=SpeciesR), shape=21, size=3, color="white") +
    geom_text(data=obj.mean[obj.mean$score=="species" & obj.mean$occur>nrow(volhybn)*minprop,], aes(label=label), color="black",  alpha=1, size=3.8, fontface=2) 
    ) 
```

## Direction of cross PERMANOVA
```{r permanova}
hybridf.plants <- svhybn.plants$SpeciesR %in% c("K x H", "H x K")
adonis2(decostand(fvolhybn.plants[hybridf.plants,], method="total")~ SpeciesR, data = svhybn[hybridf.plants,])
```

# Variation at different levels

## Variation among infloresences within plants

To measure between-trap variation, two traps were inserted into one bag, for two bags enclosing two infloresences on one plant (four traps total). Average them to get a dataset of unique infloresences.

```{r withinplant}
bd.plantsi.inflo <-  betadisper(vegdist(decostand(fvolhybn.inflo, method="hellinger"), method="bray"), 
                           group = svhybn.inflo$plantsi)
meandist.plantsi.inflo <- bd.plantsi.inflo$group.distances %>% enframe(name="plantsi", value="centroid_dist") %>%
  left_join(count(svhybn, plantsi)) %>% filter(n>1)#plants with resamples
summary(meandist.plantsi.inflo$centroid_dist)
```

## Variation among plants in each cross type

```{r amongplant, fig.width=5, fig.height=4}
#betadisper(vegdist(decostand(volhybn, method="hellinger"), method="bray"), group = specpn) #sample level
(bd.SpeciesR <- betadisper(vegdist(decostand(fvolhybn.plants, method="hellinger"), method="bray"), 
                           group = svhybn.plants$SpeciesR))
anova(bd.SpeciesR)
TukeyHSD(bd.SpeciesR)

tibble(group=bd.SpeciesR$group, distance = bd.SpeciesR$distances) %>% 
  bind_rows(bd.plantsi.inflo$distances %>% enframe(value="distance") %>% filter(distance!=0) %>% mutate(group="within")) %>%
  mutate(group = fct_relevel(group, c(levels(svhybn$SpeciesR), "within"))) %>% 
  ggplot(aes(x=group, y=distance, color=group)) +
  geom_boxplot(outlier.shape = NA) + geom_jitter(width=0.15, height=0) + geom_vline(xintercept = 4.5) + 
  scale_color_manual(values=hybcolr, guide="none") + 
  scale_y_continuous(limits=c(0,NA), expand = expansion(c(0,0.05))) + scale_x_discrete(labels=c(hyblabs, "Within plant")) + 
  theme_classic() + theme(axis.title.x=element_blank(), axis.ticks.x=element_blank()) + labs(y="Bray-Curtis distance to centroid") 
```

